Behind Bars: Insights into the Wellbeing of U.S. Prisoners

### 2024-02-16 paper/paper.qmd
Final changes

---
title: "Behind Bars: Insights into the Wellbeing of U.S. Prisoners"
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/kaavyakalani26/prison_health.git and replication aspects are available at: https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index"
date: today
date-format: long
abstract: "This research explores the causes and trends of prisoner deaths in the United States giving insights into their wellbeing in the prisons. We examine these rates both from a comprehensive standpoint and delve into demographic breakdowns, specifically considering gender and race. The paper analyse trends revealing a notable rise in deaths due to cancer and heart diseases. While AIDS-related deaths decline, homicides and drug/alcohol-related deaths show a slight increase. Comparisons across gender highlight differences in causes of death, with males experiencing higher rates of homicide, while females show a greater propensity for suicide. The examination of racial disparities exposes varying leading causes of death, emphasizing the urgent need for targeted interventions to improve the overall wellbeing of incarcerated individuals. The trends uncovered in this paper establish a foundation for further investigations into systemic problems and healthcare disparities in correctional environments."
format: pdf
number-sections: true
bibliography: references.bib
---
\newpage

@@

# Discussion
In this paper, the issue of mortality rates among U.S. inmates is analyzed and conducted.
There are multiple datasets that have been utilized to investigate demographic differences, causes of death,
and the potential impact of prison conditions. We also looked at the impact of the COVID-19 pandemic.
The study digs deeper into a complex analysis of mortality trends, drawing from datasets that include information
on gender, race, and specific causes of death. We have explored three different categories of the issue with the
wellbeing of prisoners.

>>>>

---
title: "Behind Bars: Insights into the Wellbeing of U.S. Prisoners"
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: [https://github.com/kaavyakalani26/prison_health.git](https://github.com/kaavyakalani26/prison_health.git) and replication aspects are available at: [https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index](https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index)"
date: today
date-format: long
abstract: "This research explores the causes and trends of prisoner deaths in the United States giving insights into their wellbeing in the prisons. We examine these rates both from a comprehensive standpoint and delve into demographic breakdowns, specifically considering gender and race. The paper analyse trends revealing a notable rise in deaths due to cancer and heart diseases. While AIDS-related deaths decline, homicides and drug/alcohol-related deaths show a slight increase. Comparisons across gender highlight differences in causes of death, with males experiencing higher rates of homicide, while females show a greater propensity for suicide. The examination of racial disparities exposes varying leading causes of death, emphasizing the urgent need for targeted interventions to improve the overall wellbeing of incarcerated individuals. The trends uncovered in this paper establish a foundation for further investigations into systemic problems and healthcare disparities in correctional environments."
format: pdf
number-sections: true
bibliography: references.bib
---
\newpage

@@

# Discussion

## Learnings and Inferences

In this paper, the issue of mortality rates among U.S. inmates is analyzed and conducted.
There are multiple datasets that have been utilized to investigate demographic differences, causes of death,
and the potential impact of prison conditions. We also looked at the impact of the COVID-19 pandemic.
The study digs deeper into a complex analysis of mortality trends, drawing from datasets that include information
on gender, race, and specific causes of death. We have explored three different categories of the issue with the
wellbeing of prisoners.
